### 2014-03-17 docs/paper/paper.tex
begin implementation section

Each encoder unit is attached to a circular buffer of size 8192 % FIXME check the size
containing a list of instructions for calculating one encoded value. 
This size was chosen as a compromise between memory utilization and encoder flexibility.
Each instruction corresponds to delaying for one or more clock cycles
(in order to coordinate concurrent access to the interconnect),
followed by reading one decoded value a certain address.
The decoded value data is made available to the encoder unit three clock cycles later
(due to interconnect delays), at which time it is multiplied by an encoder weight
and added to the accumulated encoded value, initially 0.
A certain bit in the instruction field acts as a flag to indicate that the encoded value is ready to be passed on to the population unit after this
instruction finishes executing. The encoder unit then waits for all other encoders to finish in order to synchronize
access to the shared interconnect, then resets the accumulator and begins encoding the next value.

@@

Encoded values are passed from the encoder units to the population units. Population units are either one-dimensional
(functions of one variable) or two-dimensional (functions of two variables).
Both types of population unit receive input from two encoder units in each input dimension;
for each dimension, the encoded values are passed through a separate first-order filter to simulate % FIXME ...post-synaptic time constant?
and then summed. The value resulting from this operation is
truncated to the most significant 10 bits and is then used as input to a set of lookup tables,
implemented as a bank of synchronous RAM elements. SRAMs are used
because the contents of these tables are not known until simulation-time and therefore the lookups cannot be implemented as hardware logic functions.
Each lookup table contains the 12-bit fixed-point values of points sampled from the principal components shared by all populations of neurons
simulated on that population unit. For one-dimensional population units, there are seven such lookup tables, and each one contains
1024 values sampled at equally-spaced points on the open interval $(-2, 2)$. For two-dimensional population units,
there are fifteen lookup tables, each containing 1024 values sampled at equally-spaced points on a square of width 4
centered at $(0, 0)$. The resulting sample space is a 32x32 grid.
To increase the accuracy of this representation, each two-dimensional population unit performs bilinear interpolation
using the least significant five bits of each input variable to determine the distance between sample points. 
The extra resources used to perform this operation are modest
and significantly reduce the amount of memory used for two-dimensional populations while maintaining a good degree of accuracy of representation.

To simulate the noise resulting from a neural representation, a pseudorandom Gaussian noise generator is implemented in each population unit.
The output from the noise generator is passed through a programmable second-order filter to approximate the power spectrum of the noise
that would be present if individual neurons were being implemented. % FIXME fact check and possibly get a citation

@@

There is an artificial limitation on the complexity of the encoding operation due to the length of the instruction buffer that is used.
With a size of 8192, % FIXME check size
because each encoder must produce 1024 encoded values per timestep, a maximum of eight instructions, and therefore eight decoded values
(not necessarily distinct)
can be used in encoding. It is theoretically possible to implement an encoding operation that requires more decoded values to be used,
but the extra instructions used for this mean that other populations sharing that encoder will have to use fewer.
It is expected that the compiler will be able to detect this scenario and reorganize populations appropriately so as to avoid overfilling the instruction buffers.
\item dimensional restrictions -- 1D, 2D population units
% FIXME citation about why 2D is ``good enough''

As only one-dimensional and two-dimensional population units have been implemented, networks that require populations to compute functions
of three or more independent variables cannot be simulated with this design.
\item four decoded values per population -- using more will create copies of populations to accommodate extra decoders

@@

\begin{itemize}
\item VC707 board
\item resources available -- BRAM, DSP
\item implemented design -- number of population units, DV buffers
\item FPGA resources utilized; maximum theoretical design size
\item clock speed and implications of this
\item Ethernet communication for control and I/O -- maximum channel capacity in terms of DVs
\end{itemize}

>>>>

Each encoder unit is attached to a circular buffer of size 8192
containing a list of instructions for calculating one encoded value. 
This size was chosen as a compromise between memory utilization and encoder flexibility.
Each instruction corresponds to delaying for one or more clock cycles
(in order to coordinate concurrent access to the interconnect),
followed by reading one decoded value a certain address.
The decoded value data is made available to the encoder unit three clock cycles later
(due to interconnect delays), at which time it is multiplied by an encoder weight
and added to the accumulated encoded value, initially 0.
A certain bit in the instruction field acts as a flag to indicate that the encoded value is ready to be passed on to the population unit after this
instruction finishes executing. The encoder unit then waits for all other encoders to finish in order to synchronize
access to the shared interconnect, then resets the accumulator and begins encoding the next value.

@@

Encoded values are passed from the encoder units to the population units. Population units are either one-dimensional
(functions of one variable) or two-dimensional (functions of two variables).
Both types of population unit receive input from two encoder units in each input dimension;
for each dimension, the encoded values are passed through a first-order filter which performs post-synaptic filtering, % FIXME check this
and then summed. The value resulting from this operation is
truncated to the most significant 10 bits and is then used as input to a set of lookup tables,
implemented as a bank of synchronous RAM elements. SRAMs are used
because the contents of these tables are not known until simulation-time and therefore the lookups cannot be implemented as hardware logic functions.
Each lookup table contains the 12-bit fixed-point values of points sampled from the principal components shared by all populations of neurons
simulated on that population unit. For one-dimensional population units, there are seven such lookup tables, and each one contains
1024 values sampled at equally-spaced points on the open interval $(-2, 2)$. For two-dimensional population units,
there are fifteen lookup tables, each containing 1024 values sampled at equally-spaced points on a square of width 4
centered at $(0, 0)$. The resulting sample space is a 32x32 grid.
To increase the accuracy of this representation, each two-dimensional population unit performs bilinear interpolation
using the least significant five bits of each input variable to determine the distance between sample points. 
The extra resources used to perform this operation are modest
and significantly reduce the amount of memory used for two-dimensional populations while maintaining a good degree of accuracy of representation.

To simulate the noise resulting from a neural representation, a pseudorandom Gaussian noise generator is implemented in each population unit.
The output from the noise generator is passed through a pre-defined second-order filter to approximate the power spectrum of the noise
that would be present if individual neurons were being implemented. % FIXME fact check and possibly get a citation

@@

There is an artificial limitation on the complexity of the encoding operation due to the length of the instruction buffer that is used.
With a size of 8192,
because each encoder must produce 1024 encoded values per timestep, a maximum of eight instructions, and therefore eight decoded values
(not necessarily distinct)
can be used in encoding. It is theoretically possible to implement an encoding operation that requires more decoded values to be used,
but the extra instructions used for this mean that other populations sharing that encoder will have to use fewer.
It is expected that the compiler will be able to detect this scenario and reorganize populations appropriately so as to avoid overfilling the instruction buffers.
\item dimensional restrictions -- 1D, 2D population units
% FIXME citation about why 2D is ``good enough''

As only one-dimensional and two-dimensional population units have been implemented, networks that require populations to compute functions
of three or more independent variables cannot be simulated with this design.
However, the majority of existing models in Nengo do not use populations with more than two input dimensions, so with the exception of
certain very large or very complex models that require representation of high-dimensional quantities, providing only one- and two-dimensional
populations is not expected to be overly restrictive.

\item four decoded values per population -- using more will create copies of populations to accommodate extra decoders

@@

\begin{itemize}
\item VC707 board
\item resources available -- BRAM, DSP

A hardware implementation of the Nengo-RT design was produced on a Xilinx VC707 development board, which includes a modern field-programmable gate array (FPGA)
device (Xilinx XC7VX485T-2FFG1761C). This device was chosen because of its ready availability on a development platform and its large logic capacity and available resources.
In particular, it provides 37~080~Kb of on-chip block RAM and 2~800 customizable DSP slices, both of which the Nengo-RT design requires in abundance in order to support
the largest possible implementation size.

\item implemented design -- number of population units, DV buffers

\item FPGA resources utilized; maximum theoretical design size
\item clock speed and implications of this
\item Ethernet communication for control and I/O -- maximum channel capacity in terms of DVs

The VC707 development board provides a Gigabit Ethernet PHY that can be used for fast off-board communication to a PC.
A simple Ethernet transceiver was implemented as part of the Nengo-RT design to expose a programming interface to the Nengo software,
as well as to provide input and output of decoded values during simulation.

\end{itemize}
